Towards a free/open-source universal-dependency treebank for Kazakh

### 2015-06-20 2015-turklang-kazdep/paper.tex
more table cleaning
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60892 72bbbca6-d526-0410-a7d9-f06f51895060
words
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60893 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{subcaption}
\usepackage{booktabs}
\setlength{\cmidrulewidth}{\heavyrulewidth}

@@

\textcite{Lynn12}, \textcite{Atalay03}, \textcite{Oflazer03}, \textcite{DeMarneffe14}

@@

\begin{table}[htbp]
  \centering
\begin{small}
  \begin{tabular}{llrrr}
    \toprule
    \textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
    \midrule
    UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
    Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
    Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
    Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
    Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
    Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
    Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
    Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
    Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
    \midrule
    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule{3-5}
  \end{tabular}
\end{small}
  \caption{Composition of the corpus. The corpus covers a range of genres and text types from 
    free and public-domain sources.}
\end{table}

>>>>

\usepackage{subcaption}
\usepackage{booktabs}

@@

This article describes the work towards the development of a dependency treebank for Kazakh, a
Turkic language spoken in Central Asia.
Despite its status as a \emph{core} Turkic
language, little computational-linguistic research has been published on syntactic
parsing for Kazakh.
A valuable resource in the study of syntactic parsing is
a treebank---a corpus of parsed text containing gold-standard syntactic annotation.

Freely available treebanks exist for many languages, such as large languages like Finnish (x,y) 
and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}. To our knowledge only one treebank
exists for another language, Turkish \parencite{Oflazer03}, which is unfortunately not freely available.

In building our treebank we take advantage of existing work done on tokenisation,
morphological analysis and part-of-speech tagging for Kazakh. We also take a pragmatic
and iterative view of development of the treebank, in line with recent work 
on cross-linguistic parsing with universal dependencies \parencite{DeMarneffe14}.

The remainder of the paper is organised as follows. Section~\ref{sec:back} gives some
background linguistic information on Kazakh, and outlines some special challenges in
parsing Kazakh. In Section~\ref{sec:method} we describe the corpus itself and methodology
used in annotating it. Section~\ref{sec:annotation} gives a sketch of some decisions
we have made with respect to annotation guidelines, referring back to the discussion in
Section~\ref{sec:back}. For reasons of space, these guidelines are not complete, but
present a subset of guidelines which are of particular interest. A small experiment
in statistical dependency parsing using the corpus is presented in Section~\ref{sec:eval},
and in Sections~\ref{sec:future} and \ref{sec:conclusions} we give perspectives
for future work and some concluding remarks.

%\textcite{Atalay03}

@@

\begin{table}[htbp]
  \centering
\begin{small}
  \begin{tabular}{llrrr}
    \toprule
	    \textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
    \midrule
	    UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
	    Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
	    Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
	    Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
	    Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
	    Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
	    Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
	    Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
	    Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
    \midrule
	    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
  \end{tabular}
\end{small}
  \caption{Composition of the corpus. The corpus covers a range of genres and text types from 
    free and public-domain sources.}
\end{table}
